---
title: "Homework 1"
author: "Phoebe Abramowitz"
date: "2/24/2018"
output: github_document
---
#Question 2
```{r setup, include=FALSE}

imports85_column_types <- list(
  'real',
  'real',
  'character',
  'character',
  'character',
  'character',
  'character',
  'character',
  'character',
  'real',
  'real',
  'real',
  'real',
  'integer',
  'character',
  'character',
  'integer',
  'character',
  'real',
  'real',
  'real',
  'integer',
  'integer',
  'integer',
  'integer',
  'integer'
)
imports85_column_names <- c(
  'symboling',
  'normalized_losses',
  'make',
  'fuel_type',
  'aspiration',
  'num_of_doors',
  'body_style',
  'drive_wheels',
  'engine_location',
  'wheel_base',
  'length',
  'width',
  'height',
  'curb_weight',
  'engine_type',
  'num_of_cylinders',
  'engine_size',
  'fuel_system',
  'bore',
  'stroke',
  'compression_ratio',
  'horsepower',
  'peak_rpm',
  'city_mpg',
  'highway_mpg',
  'price'
)

imports85 <- read.csv('imports-85.data', col.names=imports85_column_names,
         colClasses = imports85_column_types, na.strings=c("?","NA"))
str(imports85)
library(readr)
imports85_2 <- read_csv('imports-85.data', col_names = imports85_column_names, na="?",
                        col_types = list(
                          'symboling' <- col_double(),
                          'normalized_losses' <- col_double(),
                          'make' <- col_character(),
                          'fuel_type' <- col_character(),
                          'aspiration'<- col_character(),
                          'num_of_doors'<- col_character(),
                          'body_style'<- col_character(),
                          'drive_wheels'<- col_character(),
                          'engine_location'<- col_character(),
                          'wheel_base'<- col_double(),
                          'length'<- col_double(),
                          'width'<- col_double(),
                              'height'<- col_double(),
                              'curb_weight' <- col_integer(),
                              'engine_type'<- col_character(),
                              'num_of_cylinders'<- col_character(),
                              'engine_size'<- col_integer(),
                              'fuel_system'<- col_character(),
                              'bore'<- col_double(),
                              'stroke'<- col_double(),
                              'compression_ratio'<- col_double(),
                              'horsepower'<- col_integer(),
                              'peak_rpm'<- col_integer(),
                              'city_mpg'<- col_integer(),
                              'highway_mpg'<- col_integer(),
                              'price'<- col_integer()
                             ))
                               
str(imports85_2)
```

#Question 3
a)If you don't provide a vector of column names when invoking read.csv('imports-85.data'), the entries of the first row of data become the column names.

b)If you don't provide a vector of column names when invoking read.csv('imports-85.data',header=FALSE), then the column names of the imported data are "V1","V2",...,"V26"

c)If you don't specify how missing values are codified, then the data type of any column that contains "?" will be Factor.

d) The second object is bigger becuase characters take twice as many bits as integers and we specified many types as characters.

e)If you do it as an R matrix, all the ones of the type character become NA becuase they're data types that can't be transformed into integers.


#Question 4
```{r}
hist(imports85$price, col="blue")
boxplot(imports85$horsepower,horizontal = TRUE)
bs_frequency <- table(imports85$body_style)
barplot(sort(bs_frequency,decreasing = TRUE))
stars(imports85[imports85$aspiration=='turbo',c('wheel_base','length','width',
                                                  'height','price')])
```
#Question 5
```{r}
gas_cars <- imports85[imports85$fuel_type=='gas',]
diesel_cars <- imports85[imports85$fuel_type=='diesel',]
#Mean Price of Gas Type Cars is:
mean(gas_cars[,'price'],na.rm=TRUE)
#Mean Price of Diesel Type Cars is
mean(diesel_cars[,'price'],na.rm=TRUE)
#The Make of the Car with 12 cylinders is:
imports85[imports85$num_of_cylinders=='twelve',"make"]
#The make that has the most diesel cars + number of diesel cars is
diesel <- imports85[imports85$fuel_type=='diesel',]
pivot_diesel_by_make <- (table(diesel$make))
sort(pivot_diesel_by_make,decreasing=TRUE)[1]

#Price of the Car with the Most horsepower
imports85$horsepower
max(imports85$horsepower, na.rm=TRUE)
imports85$horsepower==max(imports85$horsepower, na.rm=TRUE)
imports85[which.max(imports85$horsepower),'price' ]
imports85[,"price"][which.max(imports85$horsepower)]

#Bottom 10th percentile
bottom_city_mpg <- quantile(imports85$city_mpg,0.1)
#Top 10th percentile
top_highway_mpg <- quantile(imports85$highway_mpg,0.9)
#median in bottom
median(imports85[imports85$city_mpg<=bottom_city_mpg,'price'],na.rm=TRUE)

```
#Question 6

a) "NULL" is returned
b)mtcars[, mpg] fails to return the vector 
c)This is becuase putting mpg as the second index indicates that we're trying to retrieve the row by it's name, which should be a character in quotes to indicate that its a row name.
d) YES you can include and R list of the proper length as the column of a data frame because the data frame itself is a list.
e)When you apply as.list() to a data frame, under the heading of the name of the first column, every value from that first column is listed in the same order as the data frame, and similar lists are produced for the values of every column in the data frame. 
f) You could use data.frame and input abc[i] as an argument to get desired ith column in the new dataframe.
#Question 7
```{r}

qdat <-  na.omit(imports85[,c(
  'wheel_base',
  'length',
  'width',
  'height',
  'curb_weight',
  'engine_size',
  'bore',
  'stroke',
  'compression_ratio',
  'horsepower',
  'peak_rpm',
  'city_mpg',
  'highway_mpg',
  'price')])
library(corrplot)
cor_qdat <-  cor(qdat)

#Corellelograms
corrplot(cor_qdat,method='circle')
corrplot(cor_qdat,method='color',order="hclust")
```
These two correlograms let us observe that city mpg and highway mpg are more positively correlated with each other than any two variables. It also shows us that wheelbase,length,width,curb weight,bore,horsepower,engine size,and price are all fairly positively correlated with each other. Comparatively, those variables are all negatively correlated with mpg(both city and highway). Stroke is entirely or almost entirely uncorrelated with many variables.

#Question 8.1
```{r}
imports85_pca_prcomp <- prcomp(qdat,scale.=TRUE)
imports85_pca_prcomp

eigs <- imports85_pca_prcomp$sdev^2
eigs_perc <- 100 * eigs / sum(eigs)
eigs_cum <- cumsum(eigs_perc)

eigs_df <- data.frame(
  eigenvalue = eigs,
  percentage = eigs_perc,
  'cumulative percentage' = eigs_cum
)
print(round(eigs_df, 4), print.gap = 2)
barplot(eigs, las = 1, names.arg = paste('PC', 1:ncol(qdat)), 
        main = 'Bar-chart of eigenvalues')
```
53.9% of the variation in the data is captured by the first PC, 16.2% is captured by the second PC, and 8.8% of the variation is captured by the third PC. ~89% of the variation in data is captured by the first 3 PC.
##Question 8.2
```{r}
imports85_pca_prcomp$x

plot(imports85_pca_prcomp$x, type = "n", las = 1)
abline(h = 0, v = 0)
points(imports85_pca_prcomp$x[ ,1], imports85_pca_prcomp$x[ ,2], pch=19,
       col = "blue",main="Individual Vehicles")

plot(imports85_pca_prcomp$rotation[,1],imports85_pca_prcomp$rotation[,2],col='red',
     xlab='PC2',ylab = 'PC1',pch=19,main="Eigenvectors")
text(imports85_pca_prcomp$rotation[,1],imports85_pca_prcomp$rotation[,2],labels=colnames(qdat))
```
These plots demonstrate that way_mpg,city_mpg, then horespower,engine size, bore,curb weight,width,length,and wheel base contribute largely to PC2. This makes sense with our observations from #7, as the latter group is highly negatively correlated with the former. Similarly, peak rpm and compression ration contribute heavily to PC1.

